Ants-Review: A Protocol For Open Anonymous Peer-Reviews

### 2020-09-21 paper-v1.tex
Introduction: fixes, references.

Peer-review is a necessary and essential quality control step for scientific publications but lacks proper  incentives. Indeed, the process, which is very costly in terms of time and intellectual investment, not only is not remunerated by the journals but it is also not openly recognized by the academic community as a relevant scientific output for a researcher. Therefore, scientific dissemination is affected in the timeliness, quality and fairness. Here, to solve this issue, we propose a blockchain-based incentive system that rewards scientists for peer-reviewing other scientists’ work and that builds up trust and reputation. We designed a protocol of smart contracts called Ants-Review that allows authors to issue a bounty for open anonymous peer-reviews. If requirements are met, peer-reviews will be accepted and payed by the approver proportionally to their quality and value. To promote ethical behaviour and inclusiveness the system could implement a gamified mechanism that allows the whole community to evaluate the peer-reviews, vote for the best ones and share the incentives with the referees.

@@

From its birth in 2008 as a distributed ledger for the peer-to-peer electronic cash system Bitcoin \cite{BitcoinSatoshi}, blockchain technologies (BCTs) have spread far beyond the sole cryptocurrency domain, in particular after the implementation of a general purpose smart contract functionality introduced by Ethereum \cite{Ethereum}.
Besides a growing number of applications ranging from business, healthcare, music industry, government, identity to cite but a few, blockchain technology has recently started to catalyse the attention of the scientific community as well \cite{Bitcoin-Nature-focus,vanRossum2017-DigSci} with the promising potential of being a 'game changer' in outdated and broken scientific practices and leading towards Open Science \cite{AES}. Indeed, scholars have pointed out how the intrinsic characteristics of blockchain technology set the basis for a open science infrastructure \cite{ReviewBlockchain2019} where decisional processes are transparent and therefore more democratically accessible to all the stakeholders (researchers, reviewers, funders, taxpayers). Those are: \emph{consensus algorithm}, a deterministic computational trust that allows \emph{decentralization}, for which there are no trusted third parties; \emph{cryptographic hashing} and \emph{timestamping}, that create a digital footprint able to keep a traceable chronological record of research objects; \emph{immutability} or \emph{append-only}, for which data cannot be altered or retrieved. In particular, a 'blockchainified science'\cite{BlockchainforScience} could 'reduce waste'\cite{ReducingWaste-Lancet}, by disclosing each step in the research cycle to 'scientific self-correction' way before the final publication step and therefore help fixing the current reproducibility crisis in science.
\newline A thorny issue in the academic system that can and should be tackled by blockchain concerns the status and accreditation of peer-review, the core process of scientific validation currently facing a crisis \cite{Gropp-PeerRevStress}.
In this paper we propose a solution to the problem of reviewers recognition based on the principles of token economy and in line with the values of Open Science.

>>>>

Peer-review is a necessary and essential quality control step for scientific publications but lacks proper  incentives. Indeed, the process, which is very costly in terms of time and intellectual investment, not only is not remunerated by the journals but it is also not openly recognized by the academic community as a relevant scientific output for a researcher. Therefore, scientific dissemination is affected in the timeliness, quality and fairness. Here, to solve this issue, we propose a blockchain-based incentive system that rewards scientists for peer-reviewing other scientists’ work and that builds up trust and reputation. We designed a protocol of smart contracts called Ants-Review that allows authors to issue a bounty for open anonymous peer-reviews. If requirements are met, peer-reviews will be accepted and payed by the approver proportionally to their quality and value. To promote ethical behaviour and inclusiveness the system implements a gamified mechanism that allows the whole community to evaluate the peer-reviews, vote for the best ones and share the incentives with the referees.

@@

From its birth in 2008 as a distributed ledger for the peer-to-peer electronic cash system Bitcoin \cite{BitcoinSatoshi}, blockchain technologies (BT) have spread far beyond the sole cryptocurrency domain, in particular after the implementation of a general purpose smart contract functionality introduced by Ethereum \cite{Ethereum}.
Besides a growing number of applications ranging from business, healthcare, music industry, government, identity to cite but a few, blockchain technology has recently started to catalyse the attention of the scientific community as well \cite{Bitcoin-Nature-focus,vanRossum2017-DigSci} with the promising potential of being a 'game changer' in outdated and broken scientific practices and leading towards open science \cite{AES}. Indeed, scholars have pointed out how the intrinsic characteristics of blockchain technology set the basis for a open science infrastructure \cite{ReviewBlockchain2019} in which decisional processes are transparent and therefore more democratically accessible to all the stakeholders (researchers, reviewers, funders, taxpayers). Those are: the consensus algorithm \cite{ConsAlg}, a deterministic computational trust that allows for \emph{decentralization}, for which there are no trusted third parties; the Proof of Existence (PoE) that via \emph{cryptographic hashing} and \emph{timestamping}, creates a digital footprint able to keep a traceable chronological record of research objects that cannot be altered or retrieved (due to its property of \emph{immutability} or \emph{append-only})\cite{ReviewBlockchain2019}. In particular, a 'blockchainified science'\cite{BlockchainforScience} could 'reduce waste'\cite{ReducingWaste-Lancet}, by disclosing each step in the research cycle to 'scientific self-correction' way before the final publication step and therefore help fixing the current reproducibility crisis in science.
\newline A thorny issue in the academic system that can - and should - be tackled by blockchain concerns the status and accreditation of peer-review, the core process of scientific validation currently facing a crisis \cite{Gropp-PeerRevStress}.
In this paper we propose a solution to the problem of reviewers recognition based on the principles of token economy \cite{TokenEcon} and in line with the values of open science.
